---
title: "What Explains the Electoral Success of the AfD?"
subtitle: "An Analysis of the March 2016 State Elections in Germany"
author: "Md Mujahedul Islam & Johannes Schulz-Knappe"
date: |
  | May 16, 2016
  | Hertie School of Governance
output:
  pdf_document: default
  fonttheme: structuresmallcapsserif
  html_document: null
  colortheme: beaver
  slide_level: 2
  theme: Warsaw
header-includes: \widowpenalties 1 150

bibliography:
- Library.bib
- Packages1.bib
- Packages2.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

``` {r initialisation, echo = F, message = F, warning = F}
#options(repos=c(CRAN="http://mirrors.softliste.de/cran/"))

rm(list = ls())

## Load packages for analysis

load <- c("repmis", "magrittr", "scales", "betareg", "dplyr", "xtable", "texreg", 
          "stargazer", "readxl", "ggplot2", "grid", 
          "captioner", "foreign", "rms", "Zelig", 
          "knitr", "GGally", "gridExtra", 
          "zoo", "rugarch", "psych")

loaded <- lapply(load, function(x) {
  if(!require(x, character.only = T)) {
    install.packages(x)
    require(x, character.only = T)
  }
})

## Create bibTeX file for the used packages
LoadandCite(load, file = 'Packages2.bib')

rm(load, loaded)

figcap <- captioner(prefix = "Figure", 
                    auto_space = TRUE, levels = 1, type = NULL)
tablecap <- captioner(prefix = "Table", 
                      auto_space = TRUE, levels = 1, type = NULL)


# Set the working directory:

# Create list of commonly used working directories (update, if needed)
possible_dir <- c('C:/Users/Johannes SK/Dropbox/Studium/Spring2016/CollaborativeResearch/Final-Project',
                  'C:/Users/User/Documents/GitHub/Final-Project')

set_valid_wd(possible_dir) # Set to first valid directory in the possible_dir vector
rm(possible_dir) # remove possible_dir vector

```

```{r data, echo=F, warning=F, message=F, error=F}
load("Data.Rda") # Load data file
```

# Abstract

This paper analyses factors that lead to the recent electoral success of the newly-formed "Alternative for Germany" (AfD) party in the three electoral districts of Rhineland-Palatinate, Baden-Württemberg, Saxony-Anhalt. It uses a newly generated electoral data provided by the State Offices for Statistics and structural data on the administrative districts provided by the Federal Statistical Office DESTATIS. By employing both ordinary OLS and beta regression method using these datasets, this paper finds that increase in unemployment rate is the most evident factor that lead to the recent electoral landslide for the AfD. However, the result on interaction model shows that the effects of unemployment on AfD's success varies depending on districts and this effects are mostly driven by Saxony Anhalt. The result furthermore shows that increase in the vote share of CDU and SPD in the last election also contributed to the AfD's success. However, the result from another interaction model shows that the marginal effects of CDU's vote share on AfD's electoral success actually changes depending on the vote share of SPD.  

# Introduction

Within less than three years of being founded by discontented members of the incumbent Christian Democratic Union (CDU), the newly-formed party "Alternative for Germany" (AfD) has already achieved extraordinary vote share in the General election of 2013, the European Parliament election of 2014, [@arzheimer2015afd] and lately in the three state elections in Baden-Württemberg, Rhineland-Palatinate, and Saxony-Anhalt held on March 2016. The rise of a party that stands against European Integration, argues for a re-arrangement of Germany's liberal foreign and economic policies, and vows for taking strong measure against immigrants, has presented a novel scenario into understanding factors that drive electorate to vote for a party that seemingly stands against the 'norm of the land'. Thus this research project endeavors to understand the factors that explain the success and rise of AfD in Germany's political landscape. In the following, we present current literature, the selected datasets, our applied methodology, and our findings. All work is done using the software R [@CiteR].

## The Research Question

On March 13, 2016, Germany's political landscape experienced "a landslide"[@Zeit2016]. At the three state elections in Bade-Württemberg, Rhineland-Palatinate, and Saxony-Anhalt, the AfD received double digit results from scratch, meaning that it has not been present in the state parliaments before. In Baden-Württemberg (BW), the party received 15.1%, in Rhineland-Palatinate (RP) 12.6%, and in Saxony-Anhalt (SA) even 24.3% [@BWStateOffice; @RPStateOffice; @SAStateOffice], becoming the third strongest party in BW and RP, and the second strongest in SA.
Not only does the tremendous success of this new political player harden the setup of the future governments in the respective states, it is also a clear sign that the party has arrived in German mainstream politics [@Spiegel2016].

Our research question is:

*"What explains the recent electoral success of the AfD in the different administrative districts of Baden-Württemberg, Rhineland-Palatinate, and Saxony-Anhalt?"*

We want to understand which structural factors affect the electoral success of the AfD and whether the characteristics of its voters identified in post-election analyses are reflected in the district level data. Also, we want to find out whether it follows the same patterns as the electoral success of radical right populist parties in the rest of Western Europe.

## About the AfD

Having been founded in February 2013 initially as a eurosceptic party, the AfD just missed the entry into the German national parliament at the federal elections in September 2013 by only 125,000 votes [@schmitt2014euro]. Not a single party in the post-war era was that successful in national elections being only established seven months before [@niedermayer2015neue]. 
This emergence is attributed to the Euro crisis and chancellor Merkels strong state interventionist austerity politics that opened political potential to the right of the established right wing parties CDU/CSU and FDP. The AfD recognized that potential and positioned itself initially as an expert knowledge based eurosceptic party. But over the last years, it became a new host for now unrepresented national conservatives as well as right wing populists. The unifying factors for these different groups are the anti-establishment attitude of the AfD and the image of being the political underdog who is "honest" in face of political and medial lies. In this regard, the AfD shares similarities to other European right wing populist parties [@hausler2015rechten].

The recent refugee crisis that has hit Europe and especially Germany is seen as a major factor that led to new and increased support for the AfD. Chancellor Angela Merkels slogan "Wir schaffen das", expressing optimism with regards to the capability of government and society to handle the refugee crisis, was mostly criticized by members of the AfD and to some extend from the CDU itself. As long as the refugee crisis persists, it is assumed that the AfD is going to continue to gain more support [@Spiegel2016].After the March elections, a large debate arose within the CDU/CSU whether the refugee crisis response by chancellor Merkel was responsible for the loss in votes for the CDU. Interestingly, leaders of other parties who supported Merkels "Wir schaffen das" experienced less loss of votes. However, left parties such as "Die Linke" seemed to be especially vulnerable to the emergence of the AfD [@Spiegel2016]. Also, the social democrats (SPD) seem to have lost many voters to the AfD [@Zeit2016].

# Literature review

Early analysis of the elections has shown that the AfD was especially successful in mobilizing former non-voters [@Spiegel2016]. It was also very successful amongst laborers and unemployed people. In Saxony-Anhalt, almost one third of this group voted for the AfD. But the analysis also showed that more people with medium education and income voted for the AfD. It is suggested that they are increasingly afraid of change, see the refugee crisis as a culmination of external threat, and are skeptical in the governments capacity to solve it [@Zeit2016].

In a mixed-methods study focusing on the AfD's history, self-description, positioning in the public discourse, and its supporters, @Berbuir2015 found that the party can be described as a "functional equivalent for a right-wing populist party in a country where right-wing politics are strongly stigmatized" [@Berbuir2015]. For our research this means that the AfD can be located in the tradition of new radical right wing populist (RRWP) parties in Western Europe.

When it comes to the understanding of the rise of RRWP parties, @Ivarsflaten2008 and @swank2003globalization offer important insights.
In her large scale study comparing the electoral success of RRWP parties, @Ivarsflaten2008 found that successful RRWP parties almost always used anti-immigration rhethorics in their electoral campaigns. The other two focus areas of RRWP parties - opposing political elitism and complains about the economic development - were not always instrumentalized by successful RRWP parties and seem to be more context specific.
@swank2003globalization analyzed the impact of globalization on the electoral success of RRWP parties in national elections in 16 Western European nations between 1981 and 1998. Their core finding was that the generosity of the welfare state was an important factor that mitigated the impact of economic globalizations on the rise of RRWP parties. Besides that, they found that the volume of refugees and asylum seekers, the level of international immigration, the height of domestic tax burden, previous electoral success of left libertarian parties, and the past vote share of the RRWP parties all systematically and positively influenced their electoral success. In some occasions, previously declining strength of established right-wing parties was another factor that explained RRWP party success. Interestingly, slow economic growth and the rate of people employed in manufacturing jobs did not have a significant impact [@swank2003globalization, 230].

# Hypotheses

Based on the insights from previous research on the success of RRWP parties in Western Europe and the findings from post-election analysis for the AfD, we aim to test the following hypotheses for the three state elections.

Demographic hypotheses

> *H1a*: The average level of education in a district has a negative impact on the electoral success of the AfD.

> *H1b*: The unemployment rate in a district has a positive impact on the electoral success of the AfD.

Foreign exposure hypothesis

> *H2*: The volume of asylum seekers in a district has a positive impact on the electoral success of the AfD.

Economic hypothesis

> *H3*: Local GDP per capita has a negative impact on the electoral success of the AfD.

Political hypotheses

> *H4a*: The rate of non-voters in previous state elections in a district has a positive impact on the electoral success of the AfD.

> *H4b*: The success of left wing parties in previous state elections in a district has a positive impact on the electoral success of the AfD.


# Data Sources and Methodology

In order to test these hypotheses, we combined two main data types in order to generate our variables.

The first data type is the electoral data provided by the State Offices for Statistics on the last state elections in Baden-Württemberg [@BWStateOffice], Rhineland-Palatinate [@RPStateOffice], and Saxony-Anhalt [@SAStateOffice]. The variables retrieved from these datasets are:

Variable | Name in dataset | Description |
--------------------|-----------------|-------------------------------|
Dependent Variable (DV) | **vote.AfD** | The vote share of the Alternative for Germany in the current state election |
Independent Variable 1 (IV1) | **lag.turnout** | Overall voter turnout of the previous state election                        |
IV2, IV3, IV4, IV5, IV6 | **lag.CDU**, **lag.Greens**, **lag.SPD**, **lag.FDP**, **lag.Linke** | The vote share of political parties in the previous state election (CDU, SPD, Greens, FDP, and Linke) |

The second data type is on the structural characteristics on the administrative districts. We retrieve this data from the regional data base of the Federal Statistical Office DESTATIS [@DESTATIS]. As this data is available on district level, we chose the districts as our level of analysis. The variables retrieved from these datasets are:

Variable | Name in dataset | Description |
--------------------|-----------------|-------------------------------|
IV7 | **abitur.ratio**| Ratio of school leavers per district with general qualification for university entrance (Abitur) |
IV8 | **nodegree.ratio** | Ratio of school leavers per district with no school degree |
IV9 | **GDP.cap** | Gross domestic product (GDP) per capita (by 1000)| 
IV10 | **unemp.rate** | Unemployment rate in the district |
IV11 | **n.refugees** | Number of asylum seekers per district |
IV12 | **public.debt** | Aggregated debt of the municipalities in a district (by 1000) |
IV13 | **district.type** | Categorical variable, 1 for urban districts, 0 for country districts |

## Dataset preparation

**Election Data** - Gathering the election data was complicated due to the independence of the Statistical State Offices. Every German state office chooses its own format to provide data online. While we thought about collecting data on all state elections in Germany where the AfD has won seats, we had to limit our efforts to the three states Baden-Württemberg (BW), Rhineland-Palatinate (RP), and Saxony Anhalt (SA) in order to provide a working dataset in the given time frame.
We retrieved district level data on the 2016 and 2011 state elections from the respective websites. While the data was available in separate files for RP and SA, district level data was only available for BW in a combined table that had to be divided in following steps. In total, we gathered five different datasets on the elections.

**Structural Data** - The website of the DESTATIS online database "Genesis-Online" was the source of the structural data files. As DESTATIS does not offer an API, the data had to be manually downloaded from the database. We saved four datasets as csv files that were subsequently loaded into our project.

The biggest challenge during the data cleaning was the generation of a common identifier for the different dataset. Except for the SA data, the election datasets is not provided with the district key which is the main identifier for the DESTATIS data. In order to add the identifier to the election data, we converted the district names that were available in both the election and the structural DESTATIS data into the same format and structure. Then, we sorted the data frames in the same way to add the identifier to the election data.

Finally, we merged all datasets into a single one which contains the variables mentioned above, the election year, and a state indicator on 94 observations.

## Analysis Methods

# Empirical Findings

With the generated dataset we were able to conduct our analysis. Since our dependent variable is a [0, 1] bounded variable (meaning, the vote share can take values between 0% and 100%), we used beta regression for our analysis. In the beginning we used OLS to get a basic idea of the relationships of the selected variables.

## Descriptive Statistics
Table 3 shows general descriptive statistics of our variables. Except for the variables "GDP per capita" and "Public Debt" (both divided by 1000) and "number of refugees" (absolute numbers), they are all in percent. 

The standard deviations for "GDP per capita" and "vote share of Linke" are particularly large in comparison to the respective means. GDP per capita differs substantially between rural districts and heavily urbanized ones and while the Linke is very strong in Saxony Anhalt, it is barely present in BW and RP.

```{r sumstat, echo=F, warning=F, message=F, error=F, results='asis'}
summary_labels <- c('Vote share of AfD in 2016', 'Vote turnout in 2011', 
                    'Vote share of CDU in 2011', 'Vote share of Greens in 2011', 
                    'Vote share of SPD in 2011', 'Vote share of FDP in 2011', 
                    'Vote share of Linke in 2011', 'Abitur ratio', 
                    'No degree ratio', 'GDP per capita / 1000', 
                    'Unemployment rate', 'Number of refugees', 'Public debt / 1000')

stargazer(Data[4:17],
          covariate.labels = summary_labels,
          font.size = 'tiny',
          title= 'Summary statistics of the covariates',
          type="latex", header = FALSE)
```

## Correlations among variables
Table 4 provides a correlation matrix of a selection of our variables. 

It is important to note that the AfD vote share is negatively correlated with the SPD vote share, but only slightly with the CDU vote share. GDP per capita and high school degree ratio are negatively correlated with AfD vote share, while the unemployment rate and the number of refugees are positively correlated with it.

When it comes to IV/IV correlations, the large positive correlation between GDP per capita and the high school degree ratio is not surprising. However, it is interesting to see that the high school degree ratio is negatively correlated with the previous vote share of the CDU, and positively with the unemployment rate.

```{r corr, echo=F, warning=F, message=F, error=F, results='asis'}
correlation.matrix <- cor(Data[, c("vote.AfD","lag.CDU",
                                   "lag.SPD","GDP.capita",
                                   "unempl.rate",
                                   "abitur.ratio")])
stargazer(correlation.matrix,
          title= 'Correlation matrix of some variables',
          digits = 2,
          type="latex", 
          header = FALSE, 
          no.space = FALSE)
```       

## Distribution of AfD Vote Share
When lining up the districts according to the vote share of the AfD, it becomes clear that the party achieved an overall stable and even vote share. The only real outliers are a couple of districts in Saxony Anhalt where the AfD achieved it's by far greatest successes.

```{r bar, echo=F, warning=F, message=F, error=F, results='hide', fig.width=13}
ggplot(Data) +
  geom_bar(aes(y = vote.AfD, 
               x = reorder(district.name, -vote.AfD),
               fill = state), 
           stat = "identity") +
  ylab("AfD vote share (in %)") +
  xlab("Districts of Baden-Württemberg, Rhineland-Palatinate, and Saxony Anhalt") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
```

## Correlations

In the following, we test our hypotheses with simple correlations.

### Correlation of AfD vote share and abitur ratio (H1a)

... Add text ...

``` {r fig1, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = abitur.ratio), vote.AfD, colour = state)) + 
  xlab("Abitur ratio per district (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```

### Correlation of AfD vote share and unemployment rate (H1b)

The following graph shows a statistically significant (the t value from the Pearson's product-moment correlation test is 6.27) positive correlation of the logged unemployment rate in a district with the AfD vote share. However, the correlation seems to be negative for districts with lower unemployment rates but becomes strongly positive for districts with high unemployment. The positive correlation of the abitur ratio with the unemployment rate and its negative correlation with the AfD vote share visible in the correlation matrix could offer a possible explanation for this. We will test this later by controlling for urbanization.

``` {r fig2, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = unempl.rate), vote.AfD, colour = state)) + 
  xlab("Unemployment rate per district (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```

### Correlation of AfD vote share and number of refugees (H2)

The result shows a positive correlation of the number of refugees per district with the AfD vote share. But it is not statistically significant since the t value is 0.13 (from the Pearson's product-moment correlation test). Since the data on the refugees per district is from 2013, it is highly outdated. We were not able to acquire more recent data (for 2014 or even for 2015) in order to conduct a meaningful analysis here.

``` {r fig3, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data,aes(log(x = n.refugees), vote.AfD, colour = state)) + 
  xlab("Number of refugees per district (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_light()
```

### Correlation of AfD vote share and GDP per capita (H3)
The following graph shows a slightly statistically significant (t value is -2.04) negative correlation of GDP per capita with AfD vote share. The effect seems to be mostly influenced by a small group of outliers (high vote share, low GDP per capita) and the variance is not equal across the independent variable (posing a problem with heteroscedasticity).

``` {r fig4, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = GDP.capita), vote.AfD, colour = state)) + 
  xlab("GDP per capita per district / 1000 (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```

### Correlation of AfD vote share and voter turnout in previous elections (H4a)

... Add text here ...

``` {r fig5, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = lag.turnout), vote.AfD, colour = state)) + 
  xlab("Previous voter turnout per district (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```

### Correlation of AfD vote share and strength of left wing parties (H4b)

... Add text here ...

``` {r fig6, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = lag.SPD), vote.AfD, colour = state)) + 
  xlab("Previous voteshare of SPD per district (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```



# Methods
This paper employs two empirical methods i.e., multivariate Ordinary Least Square (OLS) regression and Beta regression to examine factors that lead to the recent electoral success of AfD. The selection of method is also theoretically driven. Since our response variable is continuous, we have chosen OLS method. The class of OLS regression models is commonly used by researchers to model variables that take interval valuses as is the case for our dependent variable. The reason, on the other hand, to choose the beta regression method is for the precision and robustness purpose. Beta regression assumes that the dependent variable is beta-distributed and that its mean is related to a set of regressors through a linear predictor with unknown coefficients and a link function. Thus the beta regression is highly pertinent for our study since we are going to analyze vote share of AfD which is scaled from 0 to 1, inclusive, although 1 (complete vote received by AfD) is virtually impossible, and in our data, we observe values only up to .2944 (29.440/100). One of the merits of this method is that it includes a precision parameter which may be constant or rely on a (potentially different) set of regressors through a link function as well. Thus this approach naturally incorporates features such as heteroskedasticity or skewness which are commonly observed in data taking values in the standard unit interval, such as rates or proportions [@cribari2009beta].


# Results

## Empirical findings from ordinary multivariate OLS regression analysis

Since the initial descriptive statistics and graphs indicate certain relations and interactions, we built several regression models which we ran as OLS regression analysis first and then as beta regression. The result from the base model (1) shows increase in per capita GDP reduces the vote share of AfD by .08 while one percentage increase in unemployment rate increases the vote share of AfD by 1.05 percentage point and these results are statistically significant. Surprising the findings show that increase in the number of refugees does not have statistically significant effect on AfD's vote share. However, our base model does not consider many important variables and thus there is a high possibility of omitted variable bias. Thus to check the robustness of the findings, this paper adds other covariates in different models shown in Table . The final model (2) controls election results variable of other parties in the previous election and it shows that vote share of SPD in 2011 election and overall voter turnout in 2011 election  significantly reduces the vote share of AfD by .20 and .24 percentage point respectively.  In model (3), we added abitur ratio (high school degree) and no degree ratio to the base model and result shows that one percentage point increase in high school degree reduces the AfD's vote share by .28 percentage point. In model (4), we controlled public debt and state variable to the base model and the result shows that relative to Rhineland-Palatinate, Baden-Württemberg has negative effects on AfD's vote share by 4.91 unit. Finally, in model (5), we added all the controls to the base model and the result shows that the effect of unemployment on increasing AfD's vote shore is consistent across all models. Besides, it shows that one percentage increase in CDU and SPD's vote share in the past election also increases the vote share of AfD though it was not the case in previous model. The result also shows that relative to Rhineland-Palatinate state, Saxony-Anhalt has positive effects on AfD's vote share. Thus based on the findings from OLS model, we can say that the recent electoral success of AfD is most likely due to the increase in unemployment rate  and also due to the increase in the CDU and SPD's vote share in the past election. Also it shows that the effect of Saxony-Anhalt is very strong on AfD's vote share. This model seems to be strong since we see that, with a parsimonious sample size with few controls, the R and adjusted R^2 is .78 which means about 78 percent of the variation in AfD's vote share is explained by our model. The findings of model (6) and (7) are discriped in the following figures. 


``` {r regout, echo = F, message = F, results = "asis", error=F}
Data$state=as.factor(Data$state)
reg1 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees, data = Data)
reg2 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees + lag.CDU + 
             lag.SPD + lag.turnout, data = Data)
reg3 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees + 
             abitur.ratio + nodegree.ratio, data = Data)
reg4 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees 
           +  public.debt + state, data = Data)
reg5 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio + public.debt + state, data = Data)
reg6 <- lm(vote.AfD ~ GDP.capita + unempl.rate * state + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio + public.debt, data = Data)
reg7 <- lm(vote.AfD ~ GDP.capita + unempl.rate +state + n.refugees 
           + lag.CDU * lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio + public.debt, data = Data)

var_labels <- c('GDP per capita / 1000', 'Unemployment rate', 'Number of refugees', 
                'Vote share of CDU in 2011 election',
                'Vote share of SPD in 2011 election',
                'Voter turnout in 2011 election',
                'Abitur ratio', 'No degree ratio', 'Public debt', 'stateBW')

stargazer::stargazer(reg1, reg2, reg3, reg4, reg5, reg6, reg7 ,
                     omit = 'as.factor*', 
                     omit.stat = c('f', 'ser'), # to nicely fits on the page
                     out.header = F,
                     title = 'Determinants of electoral success for the Alternative for Germany',
                     dep.var.labels = 'Vote share of AfD',
                     covariate.labels = var_labels,
                     label = 'AfD_voteshare',
                     add.lines = list(c('District FE', rep('NO', 5))),
                     float = F, df = F, # add float and df to F
                     type="latex", header = F, 
                     no.space = T,
                     digits = 2) # add no.space=TRUE to remove space in betwen line. 
```

The result from OLS analysis shows that unemployment has strong and consistent effects on increasing AfD's vote share. Thus this paper further more runs an interaction model unemployment rate and districts level to see whether the effects of unemployment rate on AfD's vote share varies depending on districts. The result shows that most of the effects of unemployment is driven by Saxony-Anhalt. The increase in unemployment rate up to approximately 7 percentage does not have significant effect on AfD's vote share in Rhineland-Palatinam and Baden-Wurttemberg. However, the figure shows that increase in unemployment rate has a statistically significant effect on AfD vote share in Saxony-Anhalt. Thus we find evidence that unemployment rate has state based effect on increasing AfD's vote share. 

```{r AfD.prcnt, echo=F, warning=F, message=F, error=F}

# Create new variable by dividing vote.AfD variable by 100
Data$vote.AfD.prcnt <- Data$vote.AfD/100 

# Reorder variables to replace vote.AfD.prcnt variable in place of vote.AfD
reord <- Data[c(0:3, 19, 5:18, 4)] 
Data <- reord

```

```{r interaction, echo=F, warning=F, message=F, results = "asis", error=F}

#rega1 <- lm(vote.AfD.prcnt ~ state*unempl.rate, data = Data)

Data$state=as.factor(Data$state)
rega1 <- lm(vote.AfD.prcnt ~ GDP.capita + state*unempl.rate + state + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio, data = Data)
Data$predicted=predict(rega1)
apatheme=theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(),
        text=element_text(family='Times'))

p=ggplot(Data, aes(x=unempl.rate, y=vote.AfD.prcnt, shape=state))+
geom_point()+
scale_shape_manual(values=c(1,6,9), name='Districts', labels=c('Rhineland-Palatinate','Baden-Württemberg', 'Saxony-Anhalt'))+
geom_line(aes(x = unempl.rate, y = predicted, linetype=state)) +
scale_linetype_discrete(name='Districts', labels=c('Rhineland-Palatinate','Baden-Württemberg', 'Saxony-Anhalt'))+
labs(x = 'Unemployment rate', y = 'Vote share of AfD')+
apatheme
p
#ggsave('scatter1.png', width=6, height=6, unit='in', dpi=300)
```

In model (7) of the OLS regression, this paper makes an interaction model of CDU with SPD's vote share to see whether the effects on CDU's lagged vote share on AfD's electoral success varies as SPD's vote share increases. The result from the following figure shows that the impact of 1 extra (percentage) vote of CDU on AfD's vote share reduces as the vote share of SPD rises. In other words, which means that it SPD gets more vote, any extra vote to CDU has a lower effect on AfD vote. The result also shows that up to 30 % of vote share of SPD, the marginal impact of CDU vote share on AfD's vote share significantly decreases Thus this interaction model provides evidence that the effects of CDU's vote share on AfD's vote share is not constant rather it varies on the SPD's vote share.   

             
``` {r betareanalyis, fig.width = 12, fig.height = 10, echo = F, message = F, results = "asis", error=F}
interaction_plot_continuous <- function(model, effect, moderator, 
                                        interaction, varcov="default", minimum="min", 
                                        maximum="max", incr="default", num_points = 50,
                                        conf=.95, mean=FALSE, median=FALSE, alph=80,
                                        rugplot=T, histogram=T, title="Marginal effects plot", 
                                        xlabel="Value of moderator", ylabel="Estimated marginal coefficient"){
  
  # Define a function to make colors transparent
  makeTransparent<-function(someColor, alpha=alph){
    newColor<-col2rgb(someColor)
    apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                                blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
  }
  
  # Extract Variance Covariance matrix
  if (varcov == "default"){
    covMat = vcov(model)
  }else{
    covMat = varcov
  }
  
  # Extract the data frame of the model
  mod_frame = model.frame(model)
  
  # Get coefficients of variables
  beta_1 = model$coefficients[[effect]]
  beta_3 = model$coefficients[[interaction]]
  
  # Set range of the moderator variable
  # Minimum
  if (minimum == "min"){
    min_val = min(mod_frame[[moderator]])
  }else{
    min_val = minimum
  }
  # Maximum
  if (maximum == "max"){
    max_val = max(mod_frame[[moderator]])
  }else{
    max_val = maximum
  }
  
  # Check if minimum smaller than maximum
  if (min_val > max_val){
    stop("Error: Minimum moderator value greater than maximum value.")
  }
  
  # Determine intervals between values of the moderator
  if (incr == "default"){
    increment = (max_val - min_val)/(num_points - 1)
  }else{
    increment = incr
  }
  
  # Create list of moderator values at which marginal effect is evaluated
  x_2 <- seq(from=min_val, to=max_val, by=increment)
  
  # Compute marginal effects
  delta_1 = beta_1 + beta_3*x_2
  
  # Compute variances
  var_1 = covMat[effect,effect] + (x_2^2)*covMat[interaction, interaction] + 2*x_2*covMat[effect, interaction]
  
  # Standard errors
  se_1 = sqrt(var_1)
  
  # Upper and lower confidence bounds
  z_score = qnorm(1 - ((1 - conf)/2))
  upper_bound = delta_1 + z_score*se_1
  lower_bound = delta_1 - z_score*se_1
  
  # Determine the bounds of the graphing area
  max_y = max(upper_bound)
  min_y = min(lower_bound)
  
  # Make the histogram color
  hist_col = makeTransparent("grey")
  
  # Initialize plotting window
  plot(x=c(), y=c(), ylim=c(min_y, max_y), xlim=c(min_val, max_val), xlab=xlabel, ylab=ylabel, main=title)
  
  # Plot estimated effects
  lines(y=delta_1, x=x_2)
  lines(y=upper_bound, x=x_2, lty=2)
  lines(y=lower_bound, x=x_2, lty=2)
  
  # Add a dashed horizontal line for zero
  abline(h=0, lty=3)
  
  # Add a vertical line at the mean
  if (mean){
    abline(v = mean(mod_frame[[moderator]]), lty=2, col="red")
  }
  
  # Add a vertical line at the median
  if (median){
    abline(v = median(mod_frame[[moderator]]), lty=3, col="blue")
  }
  
  # Add Rug plot
  if (rugplot){
    rug(mod_frame[[moderator]])
  }
  # Add Histogram (Histogram only plots when minimum and maximum are the min/max of the moderator)
  if (histogram & minimum=="min" & maximum=="max"){
    par(new=T)
    hist(mod_frame[[moderator]], axes=F, xlab="", ylab="",main="", border=hist_col, col=hist_col)
  }
}

#png(paste0("/home/devvart/Desktop/", "marginal.png"),
    #width = 2400, height = 1800,
    #res = 250, type = "cairo-png")

interaction_plot_continuous(reg7, "lag.CDU", "lag.SPD", "lag.CDU:lag.SPD", mean = T,
                            title = "Marginal impact of CDU vote share on AFD's vote share",
                            xlabel = "Lagged vote share of SPD",
                            ylabel = "Impact of lagged CDU vote share on AFD")
```





## Empirical findings from Beta regression analysis

This paper furthermore runs the beta regression analysis since our dependent variable is bounded within 0 to 1. The result is almost same as with the OLS findings expect the size of effects. The final mode (5) of beta regression shows that one percentage increase in unemployment rate leads to .05 percentage increase in AfD's vote share which was .67 in OLS model. However, the result from beta regression provides strong evidence that unemployment, previous vote share of CDU and SPD and Saxony Anhalt, relative to RP, has a positive and statistically significnat effects on AfD's vote share though the magnitude of effect is smaller in compare to OLS findings. This model also seems to be a good model since it explains about 76 percent of the total variation of AfD's vote share. 

``` {r betareganalysis, echo = F, message = F, results = "asis", error=F}
betalogit1 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees, data = Data)
betalogit2 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees + lag.CDU + 
             lag.SPD + lag.turnout, data = Data)
betalogit3 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees + 
             abitur.ratio + nodegree.ratio, data = Data)
betalogit4 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio, data = Data)
betalogit5 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio + state, data = Data)

var_labels <- c('GDP per capita / 1000', 'Unemployment rate', 'Number of refugees', 
                'Vote share of CDU in 2011 election',
                'Vote share of SPD in 2011 election',
                'Voter turnout in 2011 election',
                'Abitur ratio', 'No degree ratio', 'stateBW')

stargazer::stargazer(betalogit1, betalogit2, betalogit3, betalogit4, betalogit5,
                     omit = 'as.factor*', 
                     omit.stat = c('f', 'ser'), # to nicely fits on the page
                     out.header = F,
                     title = 'Beta logit regression showing determinants of electoral success for the Alternative for Germany',
                     dep.var.labels = 'Vote share of AfD',
                     covariate.labels = var_labels,
                     label = 'AfD_voteshare',
                     add.lines = list(c('District FE', rep('NO', 5))),
                     font.size = 'tiny', # make the font tiny
                     float = F, df = F, # add float and df to F
                     type="latex", header = F, 
                     no.space = T,
                     digits = 2) # add no.space=TRUE to remove space in betwen line. 
```

... What did you do here? Please shortly explain ...

``` {r betareganalyis, echo = F, message = F, results = "asis", error=F}
betalogit <- betareg(vote.AfD.prcnt ~ unempl.rate, data = Data)
betaloglog <- betareg(vote.AfD.prcnt ~ unempl.rate, data = Data, link = "loglog")


library(ggplot2)
ggplot(Data, aes(x = unempl.rate, y = vote.AfD.prcnt)) +
  geom_point(size = 4, shape = 21) +
  scale_fill_grey() +
  geom_line(aes(y = predict(betaloglog, Data),
                colour = "log-log", linetype = "log-log")) +
  geom_line(aes(y = predict(betalogit, Data), 
                colour = "logit", linetype = "logit")) +
  scale_colour_manual("", values = c("red", "blue")) +
  scale_linetype_manual("", values = c("solid", "dashed")) +
  theme_bw()
```


# Limitations and Future Research

One of the limitations of this paper is that it has a parsimonious sample size. Besides the analysis is based only on the three districts of Germany and could not examine all the districts. Furthermore, the structural data on some of the variables particularly number of refugees is not recent. Thus future research could be carried out collecting updated data from all the districts to see what drives the electoral success of AfD. 

# Conclusion

This paper examines factors that lead to the recent electoral success of AfD. One of the novel contributions of this paper is that it uses a newly generated electoral data provided by the State Offices for Statistics and structural data on the administrative districts provided by the Federal Statistical Office DESTATIS. By employing both ordinary OLS and beta regression method using these datasets, this paper finds that increase in unemployment rate is the most evident factor that lead to the recent electoral landslide for the AfD. However, the result on interaction model shows that the effects of unemployment on AfD's success varies depending on districts and this effects are mostly driven by Saxony Anhalt. The result furthermore shows that increase in the vote share of CDU and SPD in the last election also contributed to the AfD's success. However, the result from another interaction model shows that the marginal effects of CDU's vote share on AfD's electoral success actually changes depending on the vote share of SPD.  Though this paper uses a parsimonious sample size and bases its analysis only on the three districts of Germany, the empirical model seems to be strong as it explains more than 70 percent of the total variation of AfD's vote share. However, since this paper could not analyse all the districts, future research could be carried out collecting updated data from all the districts to see what drives the electoral success of AfD and also to see whether effects of unemployment, among others, on AfD's vote is consistent or not.




## References {.allowframebreaks}