---
title: "What explains the electoral success of the Alternative for Germany?"
subtitle: "An analysis of the March 2016 elections in three German states"
author: "Md Mujahedul Islam & Johannes Schulz-Knappe"
date: | 
  | May 1, 2016
  | Hertie School of Governance
output: 
  ioslides_presentation
bibliography:
- Library.bib
- Packages1.bib
- Packages2.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

``` {r initialisation, echo = F, message = F, warning = F}
options(repos=c(CRAN="http://mirrors.softliste.de/cran/"))

rm(list = ls())

load <- c("repmis", "magrittr", "dplyr", "xtable", "texreg", 
          "stargazer", "readxl", "ggplot2", "grid", 
          "captioner", "foreign", "rms", "Zelig", 
          "knitr", "GGally", "gridExtra", 
          "zoo", "rugarch")

loaded <- lapply(load, function(x) {
  if(!require(x, character.only = T)) {
    install.packages(x)
    require(x, character.only = T)
  }
})
rm(load, loaded)

figcap <- captioner(prefix = "Figure", 
                    auto_space = TRUE, levels = 1, type = NULL)
tablecap <- captioner(prefix = "Table", 
                      auto_space = TRUE, levels = 1, type = NULL)


# Set the working directory:

# Create list of commonly used working directories (update, if needed)
possible_dir <- c('C:/Users/Johannes SK/Dropbox/Studium/Spring2016/CollaborativeResearch/Final-Project',
                  'C:/Users/User/Documents/GitHub/Final-Project')

set_valid_wd(possible_dir) # Set to first valid directory in the possible_dir vector
rm(possible_dir) # remove possible_dir vector

```


```{r data, echo=F, warning=F, message=F, error=F}
load("Data.Rda") # Load data file
```

## Introduction

- Within less than three years of being founded, the **Alternative for Germany** (AfD) has achieved extraordinary vote share in the General election of 2013, and the European Parliament election of 2014 [@arzheimer2015afd]

- March 2016: **15.1%** in Baden-Württemberg, **12.6%** in Rhineland-Palatinate, **24.3%** in Saxony-Anhalt

- "a landslide"[@Zeit2016]

- The AfD has **"arrived in German mainstream politics"** [@Spiegel2016]

## Existing Research

- AfD can be described as a "**functional equivalent for a right-wing populist party** in a country where right-wing politics are strongly stigmatised" [@Berbuir2015]

- @Ivarsflaten2008 : Successful RRWP parties almost always use **anti-immigration rhethorics**

- @swank2003globalization : **Volume of refugees**, level of **immigration**, domestic **tax burden**, and strong presence of **left-wing parties** positively influences right wing-populist electoral success

- AfD was especially successful in mobilizing former **non-voters** and the **unemployed** [@Spiegel2016]


## Research Question

*"What explains the recent electoral success of the AfD in the different districts of Baden-Württemberg, Rhineland-Palatinate, and Saxony-Anhalt?"*

- What structural factors correlate with the success
- Same factors as found for other western European countries?

## Election data

*Source: @BWStateOffice, @RPStateOffice, @SAStateOffice*

Variable | Name in dataset | Description |
--------------------|-----------------|-------------------------------|
Dependent Variable (DV) | **vote.AfD** | The vote share of the Alternative for Germany in the current state election |
Independent Variable 1 (IV1) | **lag.turnout** | Overall voter turnout of the previous state election                        |
IV2, IV3, IV4, IV5, IV6 | **lag.CDU**, **lag.Greens**, **lag.SPD**, **lag.FDP**, **lag.Linke** | The vote share of political parties in the previous state election (CDU, SPD, Greens, FDP, and Linke) |

## Structural data

*Source: @DESTATIS*

Variable | Name in dataset | Description |
--------------------|-----------------|-------------------------------|
IV7 | **abitur.ratio**| Ratio of school leavers per district with general qualification for university entrance (Abitur) |
IV8 | **nodegree.ratio** | Ratio of school leavers per district with no school degree |
IV9 | **GDP.cap** | Gross domestic product (GDP) per capita | 
IV10 | **unemp.rate** | Unemployment rate in the district |
IV11 | **n.refugees** | Number of asylum seekers per district |


## Descriptive Statistics

```{r sumstat, echo=F, warning=F, message=F, error=F, results='asis'}
summary_labels <- c('Vote share of AfD in 2016', 'Vote turnout in 2011', 
                    'Vote share of CDU in 2011', 'Vote share of Greens in 2011', 
                    'Vote share of SPD in 2011', 'Vote share of FDP in 2011', 
                    'Vote share of Linke in 2011', 'High school ratio', 
                    'No degree ratio', 'GDP per capita', 
                    'Unemployment rate', 'Number of refugees')


stargazer(Data[4:16],
covariate.labels = summary_labels,
font.size = 'small',
type="latex", header = FALSE)
```

## Correlation Matrix

```{r corr, echo=F, warning=F, message=F, error=F, results='asis'}
correlation.matrix <- cor(Data[, c("vote.AfD","lag.CDU",
                                   "lag.SPD","GDP.capita",
                                   "unempl.rate","n.refugees",
                                   "abitur.ratio")])
stargazer(correlation.matrix,
          font.size = 'small',
          type="latex", header = F)
```       

## The corelation of GDP per capita and AfD's vote share

``` {r fig1, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = GDP.capita), vote.AfD)) + 
  xlab("GDP per capita per district (in Euro)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```

## The corelation of unemployment and AfD's vote share

``` {r fig2, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = unempl.rate), vote.AfD)) + 
  xlab("Unemployment rate per district (in %)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```

## The corelation of no. of refugees and AfD's vote share

``` {r fig3, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data,aes(log(x = n.refugees), vote.AfD)) + 
  xlab("Number of refugees per district") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_light()
```

## Inferential statistics

``` {r regout, echo = F, message = F, results = "asis", error=F}
reg1 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees, data = Data)
reg2 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees + lag.CDU + 
             lag.SPD + lag.turnout, data = Data)
reg3 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees + 
             abitur.ratio + nodegree.ratio, data = Data)
reg4 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio, data = Data)

var_labels <- c('GDP per capita', 'Unemployment rate', 'Number of refugees', 
                'Vote share of CDU in 2011 election',
                'Vote share of SPD in 2011 election',
                'High school ratio', 'No degree ratio', '(Intercept)')

stargazer::stargazer(reg1, reg2, reg3, reg4,
          omit = 'as.factor*', 
          omit.stat = c('f', 'ser'), # to nicely fits on the page
          out.header = F,
          title = 'Determinants of electoral success for the Alternative for Germany',
          dep.var.labels = 'Vote share of AfD',
          covariate.labels = var_labels,
          label = 'AfD_voteshare',
          add.lines = list(c('District FE?', rep('NO', 4))),
          font.size = 'small',
          type="latex", header = F)
```

## Future research

```{r bar, echo=F, warning=F, message=F, error=F, results='hide', fig.width=13}
ggplot(Data) +
  geom_bar(aes(y = vote.AfD, 
               x = reorder(district.name, -vote.AfD),
               colour = state), 
           stat = "identity") +
  ylab("AfD vote share") +
  xlab("Districts of BW, RP, and SA") +
  ggtitle("Vote share of AfD in the districts of BW, RP, and SA") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
```


- Analysis of all state elections in Germany where the AfD is present

- Additional variables, like "immigrants per district"

- Spatial analysis

## Thank you!

## References