---
title: "What Explains the Electoral Success of the AfD?"
author: "Md Mujahedul Islam & Johannes Schulz-Knappe"
date: |
  | May 2, 2016
  | Hertie School of Governance
output:
  pdf_document: default
  fonttheme: structuresmallcapsserif
  html_document: null
  colortheme: beaver
  slide_level: 2
  theme: Warsaw
header-includes: \widowpenalties 1 150
subtitle: An Analysis of the March 2016 State Elections in Germany
bibliography:
- Library.bib
- Packages1.bib
- Packages2.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

``` {r initialisation, echo = F, message = F, warning = F}
#options(repos=c(CRAN="http://mirrors.softliste.de/cran/"))

rm(list = ls())

load <- c("repmis", "magrittr", "scales", "betareg", "dplyr", "xtable", "texreg", 
          "stargazer", "readxl", "ggplot2", "grid", 
          "captioner", "foreign", "rms", "Zelig", 
          "knitr", "GGally", "gridExtra", 
          "zoo", "rugarch", "psych")

loaded <- lapply(load, function(x) {
  if(!require(x, character.only = T)) {
    install.packages(x)
    require(x, character.only = T)
  }
})
rm(load, loaded)

figcap <- captioner(prefix = "Figure", 
                    auto_space = TRUE, levels = 1, type = NULL)
tablecap <- captioner(prefix = "Table", 
                      auto_space = TRUE, levels = 1, type = NULL)


# Set the working directory:

# Create list of commonly used working directories (update, if needed)
possible_dir <- c('C:/Users/Johannes SK/Dropbox/Studium/Spring2016/CollaborativeResearch/Final-Project',
                  'C:/Users/User/Documents/GitHub/Final-Project')

set_valid_wd(possible_dir) # Set to first valid directory in the possible_dir vector
rm(possible_dir) # remove possible_dir vector
#View(Data)
#write.csv(Data, file = "afd.csv")
#write.dta(Data, file = "afd.dta")
```


```{r data, echo=F, warning=F, message=F, error=F}
load("Data.Rda") # Load data file
newvariable <- (Data[,4]/(100)) #creating new variable by dividing vote.AfD variable by 100
Data$vote.AfD.prcnt <- newvariable # adding that variable into dataframe


reod <- Data[c(0:3,17,5:16,4)] #reordering variable to replace vote.AfD.prcnt variable in place of vote.AfD
Data <- reod # adding that into dataframe
#View(Data)
```





## Introduction

- Within less than three years of being founded, the **Alternative for Germany** (AfD) has achieved extraordinary vote share in the General election of 2013 and the European Parliament election of 2014 [@arzheimer2015afd]

- March 2016: **15.1%** in Baden-Württemberg, **12.6%** in Rhineland-Palatinate, **24.3%** in Saxony-Anhalt

- "**a landslide**"[@Zeit2016]

- The AfD has **"arrived in German mainstream politics"** [@Spiegel2016]

## Existing Research

- AfD can be described as a "**functional equivalent for a right-wing populist party** in a country where right-wing politics are strongly stigmatised" [@Berbuir2015]

- Successful right-wing populist parties almost always use **anti-immigration rhethorics** [@Ivarsflaten2008]

- **Volume of refugees**, level of **immigration**, domestic **tax burden**, and strong presence of **left-wing parties** positively influences right wing-populist electoral success [@swank2003globalization]

- AfD was especially successful in mobilizing former **non-voters** and the **unemployed** [@Spiegel2016]


## Research Question

*"What explains the recent electoral success of the AfD in the different districts of Baden-Württemberg, Rhineland-Palatinate, and Saxony-Anhalt?"*

- What structural factors correlate with the success?

- Are those the same characteristics found for other western European countries?

## Election Data

*Source: State Statistical Offices of @BWStateOffice, @RPStateOffice, and @SAStateOffice*

Variable | Name in dataset | Description |
--------------------|-----------------|-------------------------------|
Dependent Variable (DV) | **vote.AfD** | The vote share of the Alternative for Germany in the current state election |
Independent Variable 1 (IV1) | **lag.turnout** | Overall voter turnout of the previous state election                        |
IV2, IV3, IV4, IV5, IV6 | **lag.CDU**, **lag.Greens**, **lag.SPD**, **lag.FDP**, **lag.Linke** | The vote share of political parties in the previous state election (CDU, SPD, Greens, FDP, and Linke) |

## Structural Data

*Source: Federal Statistical Office [@DESTATIS]*

Variable | Name in dataset | Description |
--------------------|-----------------|-------------------------------|
IV7 | **abitur.ratio**| Ratio of school leavers per district with general qualification for university entrance (Abitur) |
IV8 | **nodegree.ratio** | Ratio of school leavers per district with no school degree |
IV9 | **GDP.capita** | Gross domestic product (GDP) per capita / 1000 | 
IV10 | **unemp.rate** | Unemployment rate in the district |
IV11 | **n.refugees** | Number of asylum seekers per district |


## Descriptive Statistics

```{r sumstat, echo=F, warning=F, message=F, error=F, results='asis'}
summary_labels <- c('Vote share of AfD in 2016', 'Vote turnout in 2011', 
                    'Vote share of CDU in 2011', 'Vote share of Greens in 2011', 
                    'Vote share of SPD in 2011', 'Vote share of FDP in 2011', 
                    'Vote share of Linke in 2011', 'High school ratio', 
                    'No degree ratio', 'GDP per capita / 1000', 
                    'Unemployment rate', 'Number of refugees')

stargazer(Data[4:16],
          covariate.labels = summary_labels,
          font.size = 'tiny',
          title= 'Summary statistics of the covariates',
          type="latex", header = FALSE)
```

## Correlation Matrix

```{r corr, echo=F, warning=F, message=F, error=F, results='asis'}
correlation.matrix <- cor(Data[, c("vote.AfD","lag.CDU",
                                   "lag.SPD","GDP.capita",
                                   "unempl.rate",
                                   "abitur.ratio")])
stargazer(correlation.matrix,
          title= 'Correlation matrix of some variables',
          font.size = 'tiny',
          digits = 2,
          type="latex", 
          header = FALSE, 
          no.space = FALSE)
```       

## Distribution of AfD Vote Share

```{r bar, echo=F, warning=F, message=F, error=F, results='hide', fig.width=13}
ggplot(Data) +
  geom_bar(aes(y = vote.AfD, 
               x = reorder(district.name, -vote.AfD),
               fill = state), 
           stat = "identity") +
  ylab("AfD vote share (in %)") +
  xlab("Districts of Baden-Württemberg, Rhineland-Palatinate, and Saxony Anhalt") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
```


## The Correlation of GDP per Capita and AfD Vote Share

``` {r fig1, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = GDP.capita), vote.AfD)) + 
  xlab("GDP per capita per district / 1000 (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```

## The Correlation of Unemployment and AfD Vote Share

``` {r fig2, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data, aes(log(x = unempl.rate), vote.AfD, colour = state)) + 
  xlab("Unemployment rate per district (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_bw()
```

## The Correlation of Number of Refugees and AfD Vote Share

``` {r fig3, echo = F, message = F, results = "hide", error=F, fig.height=4}
ggplot2::ggplot(Data,aes(log(x = n.refugees), vote.AfD)) + 
  xlab("Number of refugees per district (log)") +
  ylab("AfD vote share (in %)") +
  geom_point(aes(colour = state)) + 
  geom_smooth() + 
  theme_light()
```

## Inferential Statistics: normal multivariate OLS regression analysis

``` {r regout, echo = F, message = F, results = "asis", error=F}
reg1 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees, data = Data)
reg2 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees + lag.CDU + 
             lag.SPD + lag.turnout, data = Data)
reg3 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees + 
             abitur.ratio + nodegree.ratio, data = Data)
reg4 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio, data = Data)
reg5 <- lm(vote.AfD ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio + state, data = Data)

var_labels <- c('GDP per capita / 1000', 'Unemployment rate', 'Number of refugees', 
                'Vote share of CDU in 2011 election',
                'Vote share of SPD in 2011 election',
                'Voter turnout in 2011 election',
                'Abitur ratio', 'No degree ratio', 'state')

stargazer::stargazer(reg1, reg2, reg3, reg4, reg5,
                     omit = 'as.factor*', 
                     omit.stat = c('f', 'ser'), # to nicely fits on the page
                     out.header = F,
                     title = 'Determinants of electoral success for the Alternative for Germany',
                     dep.var.labels = 'Vote share of AfD',
                     covariate.labels = var_labels,
                     label = 'AfD_voteshare',
                     add.lines = list(c('District FE', rep('NO', 5))),
                     font.size = 'tiny', # make the font tiny
                     float = F, df = F, # add float and df to F
                     type="latex", header = F, 
                     no.space = T,
                     digits = 2) # add no.space=TRUE to remove space in betwen line. 
```


```{r interaction, echo=F, warning=F, message=F, error=F}

#rega1 <- lm(vote.AfD.prcnt ~ state*unempl.rate, data = Data)

Data$state=as.factor(Data$state)
rega1 <- lm(vote.AfD.prcnt ~ GDP.capita + state*unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio, data = Data)
summary(rega1)
Data$predicted=predict(rega1)
apatheme=theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(),
        text=element_text(family='Times'))

p=ggplot(Data, aes(x=unempl.rate, y=vote.AfD.prcnt, shape=state))+
geom_point()+
scale_shape_manual(values=c(1,6,9), name='Districts', labels=c('Rhineland-Palatinate','Baden-Württemberg', 'Saxony-Anhalt'))+
geom_line(aes(x = unempl.rate, y = predicted, linetype=state)) +
scale_linetype_discrete(name='Districts', labels=c('Rhineland-Palatinate','Baden-Württemberg', 'Saxony-Anhalt'))+
labs(x = 'Unemployment rate', y = 'Vote share of AfD')+
apatheme
p
#ggsave('scatter1.png', width=6, height=6, unit='in', dpi=300)
```




```



## Inferential Statistics: Beta regression analysis

``` {r betareganalysis, echo = F, message = F, results = "asis", error=F}
betalogit1 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees, data = Data)
betalogit2 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees + lag.CDU + 
             lag.SPD + lag.turnout, data = Data)
betalogit3 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees + 
             abitur.ratio + nodegree.ratio, data = Data)
betalogit4 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio, data = Data)
betalogit5 <- betareg(vote.AfD.prcnt ~ GDP.capita + unempl.rate + n.refugees 
           + lag.CDU + lag.SPD + lag.turnout + abitur.ratio + 
             nodegree.ratio + state, data = Data)

var_labels <- c('GDP per capita / 1000', 'Unemployment rate', 'Number of refugees', 
                'Vote share of CDU in 2011 election',
                'Vote share of SPD in 2011 election',
                'Voter turnout in 2011 election',
                'Abitur ratio', 'No degree ratio', 'state')

stargazer::stargazer(betalogit1, betalogit2, betalogit3, betalogit4, betalogit5,
                     omit = 'as.factor*', 
                     omit.stat = c('f', 'ser'), # to nicely fits on the page
                     out.header = F,
                     title = 'Beta logit regression showing determinants of electoral success for the Alternative for Germany',
                     dep.var.labels = 'Vote share of AfD',
                     covariate.labels = var_labels,
                     label = 'AfD_voteshare',
                     add.lines = list(c('District FE', rep('NO', 5))),
                     font.size = 'tiny', # make the font tiny
                     float = F, df = F, # add float and df to F
                     type="latex", header = F, 
                     no.space = T,
                     digits = 2) # add no.space=TRUE to remove space in betwen line. 
```









``` {r betareganalyis, echo = F, message = F, results = "asis", error=F}
betalogit <- betareg(vote.AfD.prcnt ~ unempl.rate, data = Data)
betaloglog <- betareg(vote.AfD.prcnt ~ unempl.rate, data = Data, link = "loglog")


library(ggplot2)
ggplot(Data, aes(x = unempl.rate, y = vote.AfD.prcnt)) +
  geom_point(size = 4, shape = 21) +
  scale_fill_grey() +
  geom_line(aes(y = predict(betaloglog, Data),
                colour = "log-log", linetype = "log-log")) +
  geom_line(aes(y = predict(betalogit, Data), 
                colour = "logit", linetype = "logit")) +
  scale_colour_manual("", values = c("red", "blue")) +
  scale_linetype_manual("", values = c("solid", "dashed")) +
  theme_bw()
```


View("Data")

## Limitations and Future Research

- Extend analysis on all state elections in Germany where the AfD has participated in order to increase observations

- Access more recent data on refugee numbers
 
- Include additional variables, like "immigration"

- Create interaction models 

- Use beta regression as dependent variable is [0, 1] bounded

## Q&A

**Thank you for your attention!**

## References {.allowframebreaks}